DP problems of interest

### 2012-09-30 docs/roadmap/problems.tex
More cleanup of problems.

% LaTeX Template for a short article
\documentclass[11pt]{article}
\usepackage{amssymb,amsmath,amsthm,hyperref,verbatim,pict2e,graphicx,marvosym,array,booktabs}
\hypersetup{colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}
\usepackage[usenames]{xcolor} % color names ,dvipsnames,svgnames,table
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\pagestyle{empty} % remove page numbers

@@

% Recurrence visualisation
\newcommand\Cd[3][0,-1]{\put(#2){\put(.5,.5){\circle*{.3}}\put(.5,.5){\linethickness{1.5pt}\vector(#1){#3}}}} % dependency [dx,dy]{x,y}{len}
\def\Cg#1{\put(#1){\color{lightgray}\put(0,0){\polygon*(0,0)(0,1)(1,1)(1,0)}}} % grayed cell (not to store
\def\Cz#1{\put(#1){\put(0,.35){\parbox{1\unitlength}{\centering\bf 0}}}} % zero-init cell
\def\Cm{\put(6.5,1.5){\circle*{.4}}\multiput(0,0)(1,0){9}{\line(0,1){8}}\multiput(0,0)(0,1){9}{\line(1,0){8}}} % matrix base

@@

% ----------------------------------------------
\newpage
\subsection{Smith-Waterman (simple)}
Problem: matching two strings $S$, $T$ with $|S|=m, |T|=n$.

Matrices: $M_{m \times n}, B_{m \times n}$

Alphabets:\ul
\item Input: $\Sigma(S)=\Sigma(T)=\{a,c,g,t\}$.
\item Cost matrix: $\Sigma(M) = [0..z], z=\max({\rm cost(\_)}) \cdot \min(m,n)$
\item Backtrack matrix: $\Sigma(B)=\{stop,W,N,NW\}$
\ule

Initialization:\ul
\item Cost matrix: $M_{(i,0)}=M_{(0,j)}=0$.
\item Backtrack matrix: $B_{(i,0)}=B_{(0,j)}=stop$.
\ule

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	0 & stop\\
	M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
	M_{(i-1,j)}-d & N\\
	M_{(i,j-1)}-d & W
\end{array}\right\}=B_{(i,j)} \]

Backtracking: starts from the cell $M_{(m,j)} \cup M_{(i,n)} $

Visualisation:
\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(8,9)
	\put(-.5,7.5){S}\put(-.35,7.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){T}\put(.8,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cz{0,0}\Cz{0,1}\Cz{0,2}\Cz{0,3}\Cz{0,4}\Cz{0,5}\Cz{0,6}\Cz{0,7}
	\Cz{1,7}\Cz{2,7}\Cz{3,7}\Cz{4,7}\Cz{5,7}\Cz{6,7}\Cz{7,7}
	\Cd{6,2}{0.8}
	\Cd[1,0]{5,1}{0.8}
	\Cd[1,-1]{5,2}{0.8}
\Cm\end{picture}\end{center}

Optimizations\ul
\item In serial (monadic) problems we can avoid building the matrix $M$ by only maintaining the 3 last diagonals in memory (one for the diagonal element, one for horizontal/vertical, and one being built). This construction extends easily to polyadic problems where we need to maintain $k+2$ diagonals in memory where $k$ is the maximum backward lookup.
\ule

% ----------------------------------------------
\newpage
\subsection{Smith-Waterman (with gap extension at different cost)}
Problem: matching two strings $S$, $T$ with $|S|=m, |T|=n$.

Matrices: $M_{m \times n}, E_{m \times n}, F_{m \times n}, B_{m \times n}$

Alphabets:\ul
\item Input: $\Sigma(S)=\Sigma(T)=\{a,c,g,t\}$.
\item Cost matrices: $\Sigma(M) = \Sigma(E) = \Sigma(F) = [0..z], z=\max({\rm cost(\_)}) \cdot \min(m,n)$
\item Backtrack matrix: $\Sigma(B)=\{stop,W,N,NW\}$
\ule

Initialization:\ul
\item No gap cost matrix: $M_{(i,0)}=M_{(0,j)}=0$.
\item T-gap extension cost matrix: $E_{(i,0)}= 0$ \textit{<<eat S chars only>>}
\item S-gap extension cost matrix: $F_{(0,j)}= 0$
\item Backtrack matrix: $B_{(i,0)}=B_{(0,j)}=stop$.
\ule

Recurrence for the cost matrices:
\[\begin{array}{rcl}
M_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	0 & stop\\
	M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
	E_{(i,j)} & N\\
	F_{(i,j)} & W
\end{array}\right\}=B_{(i,j)}\\
\\
E_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	M_{(i, j-1)} - \alpha & NW\\
	E_{(i,j-1)} - \beta & N\\
\end{array}\right\}=B_{(i,j)}\\
\\
F_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	M_{(i-1,j)} - \alpha & NW\\
	F_{(i-1,j)} - \beta & W\\
\end{array}\right\}=B_{(i,j)}
\end{array}\]

{\color{red} XXX: Are we sure we do not need to maintain separate backtrack tables for the gap extensions?}

Otherwise written as:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	0 & stop\\
	M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
	\max_{1 \le k \le j-1} M_{(i,k)} - \alpha - (j-1-k) \cdot \beta & N\\
	\max_{1 \le k \le i-1} M_{(k,j)} - \alpha - (i-1-k) \cdot \beta & W\\
\end{array}\right\}=B_{(i,j)} \]

{\color{red} XXX: how to simplify this relation, can we move the max in the wavefront?}

Recurrence visualisation:

{\color{red} XXX: todo}

Optimizations:

{\color{red} XXX: todo}

% ----------------------------------------------
\newpage
\subsection{Convex polygon triangulation}
Problem: triangulating a polygon with least cost overall for added edges

Alphabets:\ul
\item Input: $\Sigma(S)= $ a matrix of costs for every possible edge in the polygon. The 
size of the input is $\frac{n^2}{2}$, where $n$ is the number of vertices in the polygon. 
\item Cost matrix: $\Sigma(M)=\mathbb{N}$. The sum of cost of edges is not "bounded".
\item Backtrack matrix: $\Sigma(B)=\{n\}$
\ule

Recurrence:
\[M_{(i,j)}= \max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} + c(i,k) | k = B_{(i,j)} \]

It is interesting to note that even in the sequential world, this problem is best solved 
by filling the diagonals, ie. computing sub-solutions for all polygons on size $k$ before
those of size $k+1$.

% ----------------------------------------------

% ----------------------------------------------
\newpage
\subsection{Matrix chain multiplication}
Problem: find an optimal parenthesizing of the multiplication of $n$ matrices $A_i$ of dimension $r_i \times c_i$ \textit{(r=rows, c=columns)}.

Alphabets:\ul
\item Input: matrix size is defined as pairs of integers $(r_i,c_i)$.
\item Cost matrix: the cost is an integer (might blow up, use a float or a double?).
\item Backtrack matrix: $\Sigma(B)=\{stop\} \cup \{0..n\}$ with $n$ the input length.
\ule

Initialization (line=length of sequence, column=start):\ul
\item Cost matrix: $M_{(0,j)}=0$.
\item Backtrack matrix: $B_{(0,j)}=stop$.
\ule

Recurrence for the cost matrix:
\[M_{(i,j)}=\min_{i\le k<j}\left\{\begin{array}{l|l}
	M_{(i,k)}+M_{(k,j)}+r_i \cdot r_k \cdot col_j & k
\end{array}\right\}=B_{(i,j)} \]

% ----------------------------------------------
\subsection{Nussinov algorithm}
Problem: folding a RNA string $S$ over itself.

Alphabets:\ul
\item Input: $\Sigma(S)=\{A,C,G,U\}$.
\item Cost matrix: $\Sigma(M)=\{0..n\}, n={\rm length}(S)/2$
\item Backtrack matrix: $\Sigma(B)=\{stop,W,S,SW, 1..n\}$
\ule

Initialization: \ul
\item Cost matrix: $\left\{\begin{array}{l} M_{(i,i)}=0 \\ M_{(i,i-1)}=0 \end{array}\right. \forall i \in 1..{\rm length}(S)$
\item Backtrack matrix: $\left\{\begin{array}{l} B_{(i,i)}=stop \\ B_{(i,i-1)}=stop \end{array}\right.  \forall i \in 1..{\rm length}(S)$
\ule

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	M_{(i+1,j-1)}+\omega(i,j) & SW\\
	M_{(i+1,j)} & S\\
	M_{(i,j-1)} & W\\
	\max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} & k
\end{array}\right\} = B_{(i,j)} \]
With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Zuker folding}
Problem: folding a RNA string $S$ over itself.

Alphabets:\ul
\item Input: $\Sigma(S)=\{A,C,G,U\}$.
\item Cost matrix: $\Sigma(W)=\Sigma(V)=\Sigma(F)=\{0..n\},n=$
\item Backtrack matrix: $\Sigma(B)=\{\}$
\ule

Initialization:\ul
\item Cost matrices:\ul
	\item XXX: $W$,$V$
	\item $F_{(0)}=0$
\ule
\item Backtrack matrix: $B$
\ule

Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l|l}
	W_{(i+1,j)}+b & S\\
	W_{(i,j-1)}+b & W\\
	V_{(i,j)}+\delta(S_i,S_j) & ?V? \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)} & ?k?
\end{array}\right.\\
\\
V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\
\\
 F_{(j)}&=&\min\left\{\begin{array}{ll} F_{(j-1)} \\ \min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
%
\end{array}\right. \text{(Free Energy)}\\\\
VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
\end{array}\]
%

%%\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash}m{#1}}
%\newcolumntype{C}{@{\hspace{7pt}}c@{\hspace{7pt}}}
%\def\mnl{\rule{0pt}{2.6ex}\rule[-1.2ex]{0pt}{0pt} \\ \hline}
%$\begin{array}{|C|C|C|C|C|C|} \hline
%0 & 0 & 0 & 0 & 0 & 0 \mnl
%0 &  &  &  &  & \mnl
%0 & M_{23}  &  &  &  & \mnl
%0 &  & \sum  &  &  & \mnl
%0 &  &  &  &  & \mnl
%0 &  &  &  &  & \mnl
%\end{array}$

\end{document}

>>>>

% LaTeX Template for a short article
\documentclass[11pt]{article}
\usepackage{amssymb,amsmath,amsthm,hyperref,verbatim,pict2e,graphicx,marvosym,array,booktabs,nameref}
\hypersetup{colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}
\usepackage[usenames]{xcolor} % color names ,dvipsnames,svgnames,table
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\pagestyle{empty} % remove page numbers

@@

% Recurrence visualisation
\newcommand\Cd[3][0,-1]{\put(#2){\put(.5,.5){\circle*{.3}}\put(.5,.5){\linethickness{1.5pt}\vector(#1){#3}}}} % dependency [dx,dy]{x,y}{len}
\def\Cg#1{\put(#1){\color{lightgray}\put(0,0){\polygon*(0,0)(0,1)(1,1)(1,0)}}} % grayed cell (not to store
\def\Cz#1{\put(#1){\put(0,.35){\parbox{1\unitlength}{\centering\bf 0}}}} % zero-init cell
\def\Cm{\put(6.5,4.5){\circle*{.4}}\multiput(0,0)(1,0){9}{\line(0,1){8}}\multiput(0,0)(0,1){9}{\line(1,0){8}}} % matrix base

@@

% ----------------------------------------------
\newpage
\subsection{Smith-Waterman (simple)}\label{sswat}\ol
\item Problem: matching two strings $S$, $T$ with $|S|=m, |T|=n$.
\item Matrices: $M_{m \times n}, B_{m \times n}$
\item Alphabets:\ul
	\item Input: $\Sigma(S)=\Sigma(T)=\{a,c,g,t\}$.
	\item Cost matrix: $\Sigma(M) = [0..z], z=\max({\rm cost(\_)}) \cdot \min(m,n)$
	\item Backtrack matrix: $\Sigma(B)=\{stop,W,N,NW\}$
	\ule
\item Initialization:\ul
	\item Cost matrix: $M_{(i,0)}=M_{(0,j)}=0$.
	\item Backtrack matrix: $B_{(i,0)}=B_{(0,j)}=stop$.
	\ule
\item Recurrence: \[M_{(i,j)}=\max\left\{\begin{array}{l|l}
		0 & stop\\
		M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
		M_{(i-1,j)}-d & N\\
		M_{(i,j-1)}-d & W
	\end{array}\right\}=B_{(i,j)} \]

\item Backtracking: starts from the cell $M_{(m,j)} \cup M_{(i,n)} $
\item Visualisation: by convention, we put the longest string vertically ($m\ge n$):
\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(8,9)
	\put(-.5,7.5){S}\put(-.35,7.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){T}\put(.8,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cz{0,0}\Cz{0,1}\Cz{0,2}\Cz{0,3}\Cz{0,4}\Cz{0,5}\Cz{0,6}\Cz{0,7}
	\Cz{1,7}\Cz{2,7}\Cz{3,7}\Cz{4,7}\Cz{5,7}\Cz{6,7}\Cz{7,7}
	\Cd{6,5}{0.8}
	\Cd[1,0]{5,4}{0.8}
	\Cd[1,-1]{5,5}{0.8}
\Cm\end{picture}\end{center}

\item Optimizations:\ul
	\item In serial (monadic) problems we can avoid building the matrix $M$ by only maintaining the 3 last diagonals in memory (one for the diagonal element, one for horizontal/vertical, and one being built). This construction extends easily to polyadic problems where we need to maintain $k+2$ diagonals in memory where $k$ is the maximum backward lookup.
	\ule
\ole

% ----------------------------------------------
\newpage
\subsection{Smith-Waterman (with gap extension at different cost)}\ol
\item Problem: matching two strings $S$, $T$ with $|S|=m, |T|=n$.
\item Matrices: $M_{m \times n}, E_{m \times n}, F_{m \times n}, B_{m \times n}$
\item Alphabets:\ul
	\item Input: $\Sigma(S)=\Sigma(T)=\{a,c,g,t\}$.
	\item Cost matrices: $\Sigma(M) = \Sigma(E) = \Sigma(F) = [0..z], z=\max({\rm cost(\_)}) \cdot \min(m,n)$
	\item Backtrack matrix: $\Sigma(B)=\{stop,W,N,NW\}$
	\ule
\item Initialization:\ul
	\item No gap cost matrix: $M_{(i,0)}=M_{(0,j)}=0$.
	\item T-gap extension cost matrix: $E_{(i,0)}= 0$ \textit{<<eat S chars only>>}
	\item S-gap extension cost matrix: $F_{(0,j)}= 0$
	\item Backtrack matrix: $B_{(i,0)}=B_{(0,j)}=stop$.
	\ule
\item Recurrence for the cost matrices:
\[\begin{array}{rcl}
M_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	0 & stop\\
	M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
	E_{(i,j)} & N\\
	F_{(i,j)} & W
\end{array}\right\}=B_{(i,j)}\\
\\
E_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	M_{(i, j-1)} - \alpha & NW\\
	E_{(i,j-1)} - \beta & N\\
\end{array}\right\}=B_{(i,j)}\\
\\
F_{(i,j)}&=&\max\left\{\begin{array}{l|l}
	M_{(i-1,j)} - \alpha & NW\\
	F_{(i-1,j)} - \beta & W\\
\end{array}\right\}=B_{(i,j)}
\end{array}\]

That can be written alternatively as:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	0 & stop\\
	M_{(i-1,j-1)}+{\rm cost}(S(i),T(j)) & NW\\
	\max_{1 \le k \le j-1} M_{(i,k)} - \alpha - (j-1-k) \cdot \beta & N\\
	\max_{1 \le k \le i-1} M_{(k,j)} - \alpha - (i-1-k) \cdot \beta & W\\
\end{array}\right\}=B_{(i,j)} \]

Although the latter notation seems more explicit, it introduces non-serial dependencies that the former set of recurrences is free of. So we need to implement the former rules whose kernel is 
\[ [M;E;F]_{(i,j)} = f_{\rm kernel} ( [M;E]_{(i,j-1)}, [M;F]_{(i-1,j)}, M_{(i-1,j-1)} ) \]
Notice that this recurrence is very similar to \nameref{sswat} except that we propagate 3 values ($M,E,F$) instead of a single one ($M$).

\item Backtracking: same as \nameref{sswat}
\item Visualisation: same as \nameref{sswat}
\item Optimizations: same as \nameref{sswat}
\ole

% ----------------------------------------------
\newpage
\subsection{Convex polygon triangulation}\ol
\item Problem: triangulating a polygon of $n$ vertices with least total cost for added edges. We denote the cost of adding an edge between the  pair of edges $i,j$ by $S(i,j)$, Where $S_{n \times n}$ is a lower triangular matrix compacted in memory (rows are contiguous) with a 0 diagonal that is omitted \footnote{Arbitrary convention for both architectural implementation and code generator. Rationale: in lower triangular matrix, element address is independent of the matrix size.}, hence $|S|=\tfrac{n^2}{2}=N$.

% ------------------------------------------------------------------------------------------
% ------------------------------------------------------------------------------------------
% ------------------------------------------------------------------------------------------
{\color{red}
\item Matrices: $M_{n\times n}, B_{n \times n}$ \textit{<<first edge, last edge>>} upper triangular including main diagonal 
\item Alphabets:\ul
	\item Input: $\Sigma(S_{(i,j)})=\{0..m\}$ with $m=\max_S(i,j) \forall i,j$ determined at runtime\footnote{We need to scan/have stats about $S$ and that's where LMS plays a role}.
	\item Cost matrix: $\Sigma(M)=\{0..z\}$ with $z = m \cdot (n-2)$ (we add at most $n-2$ edges).
	\item Backtrack matrix: $\Sigma(B)=\{stop, 0..n\}$ (the index of the edge we add)
	\ule
\item Initialization: $M_{(i,i)}=0, B_{(i,i)}=stop \quad\forall i$
\item Recurrence: \[M_{(i,j)}=\left\{ \max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} + S(i,k) \,\,\rule[-.75em]{.5pt}{2em}\,\,  k \right\} = B_{(i,j)} \]
	It is interesting to note that even in the sequential world, this problem is best solved 
	by filling the diagonals, ie. computing sub-solutions for all polygons on size $k$ before
	those of size $k+1$.
\item Backtracking: Starting at $B_{(0,n)}$ and let current point you backtrack drawing an edge from the current position to the backtracking position ??

\item Visualisation: in its traditional form the matrix is upper triangular with diagonal excluded.
\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(8,9)
	\put(-.7,6.5){\rotatebox{90}{First}}\put(-.4,6.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){Last}\put(1.5,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cg{0,6}
	\Cg{0,5} \Cg{1,5}
	\Cg{0,4} \Cg{1,4} \Cg{2,4}
	\Cg{0,3} \Cg{1,3} \Cg{2,3} \Cg{3,3}
	\Cg{0,2} \Cg{1,2} \Cg{2,2} \Cg{3,2} \Cg{4,2}
	\Cg{0,1} \Cg{1,1} \Cg{2,1} \Cg{3,1} \Cg{4,1} \Cg{5,1}
	\Cg{0,0} \Cg{1,0} \Cg{2,0} \Cg{3,0} \Cg{4,0} \Cg{5,0} \Cg{6,0}

	\Cz{0,7}\Cz{1,6}\Cz{2,5}\Cz{3,4}\Cz{4,3}\Cz{5,2}\Cz{6,1}\Cz{7,0}
	\Cd[0,1]{6,1}{2.8}\Cd[0,1]{6,2}{1.8}\Cd[0,1]{6,3}{0.8}
	\Cd[1,0]{3,4}{2.8}\Cd[1,0]{4,4}{1.8}\Cd[1,0]{5,4}{0.8}
	\put(3.5,4.5){\line(3,-1){3}}
	\put(4.5,4.5){\line(2,-2){2}}
	\put(5.5,4.5){\line(1,-3){1}}

	%\Cd[1,-1]{5,2}{0.8}

\Cm\end{picture}\end{center}

\item Optimizations: XXX we need to rotate that matrix to progress in the same direction as usual, that is towards bottom right.
}
\ole

% ----------------------------------------------
\newpage
\subsection{Matrix chain multiplication}\ol
\item Problem: find an optimal parenthesizing of the multiplication of $n$ matrices $A_i$. Each matrix $A_i$ is of dimension $r_i \times c_i$ \textit{<<r=rows, c=columns>>}. 
\item Matrices: $M_{n \times n}, B_{n \times n}$ \textit{(first, last matrix)}
\item Alphabets:\ul
	\item Input: matrix $A_i$ size is defined as pairs of integers $(r_i,c_i)$.
	\item Cost matrix: $\Sigma(M):=$ huge integer\footnote{Integer multiplication might blow up, use float or doubles and addition of logarithms instead.}.
	\item Backtrack matrix: $\Sigma(B)=\{stop\} \cup \{0..n\}$.
	\ule
\item Initialization:\ul
	\item Cost matrix: $M_{(i,i)}=0$.
	\item Backtrack matrix: $B_{(i,i)}=stop$.
	\ule

{\color{red}
\item Recurrence for the cost matrix:
	\[M_{(i,j)}=\min_{i\le k<j}\left\{\begin{array}{l|l}
		M_{(i,k)}+M_{(k,j)}+r_i \cdot r_k \cdot col_j & k
	\end{array}\right\}=B_{(i,j)} \]
\item Backtracking: XXX
\item Visualisation: XXX

\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(8,9)
	\put(-.7,6.5){\rotatebox{90}{First}}\put(-.4,6.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){Last}\put(1.5,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cg{0,6}
	\Cg{0,5} \Cg{1,5}
	\Cg{0,4} \Cg{1,4} \Cg{2,4}
	\Cg{0,3} \Cg{1,3} \Cg{2,3} \Cg{3,3}
	\Cg{0,2} \Cg{1,2} \Cg{2,2} \Cg{3,2} \Cg{4,2}
	\Cg{0,1} \Cg{1,1} \Cg{2,1} \Cg{3,1} \Cg{4,1} \Cg{5,1}
	\Cg{0,0} \Cg{1,0} \Cg{2,0} \Cg{3,0} \Cg{4,0} \Cg{5,0} \Cg{6,0}

	\Cz{0,7}\Cz{1,6}\Cz{2,5}\Cz{3,4}\Cz{4,3}\Cz{5,2}\Cz{6,1}\Cz{7,0}
	\Cd[0,1]{6,1}{2.8}\Cd[0,1]{6,2}{1.8}\Cd[0,1]{6,3}{0.8}
	\Cd[1,0]{3,4}{2.8}\Cd[1,0]{4,4}{1.8}\Cd[1,0]{5,4}{0.8}
	\put(3.5,4.5){\line(3,-1){3}}
	\put(4.5,4.5){\line(2,-2){2}}
	\put(5.5,4.5){\line(1,-3){1}}

	%\Cd[1,-1]{5,2}{0.8}

\Cm\end{picture}\end{center}

\item Optimizations: XXX
}
\ole

% ----------------------------------------------
\newpage
\subsection{Nussinov algorithm}\ol
\item Problem: folding a RNA string $S$ over itself.
{\color{red}
\item Matrices: XXX
\item Alphabets:\ul
	\item Input: $\Sigma(S)=\{A,C,G,U\}$.
	\item Cost matrix: $\Sigma(M)=\{0..n\}, n={\rm length}(S)/2$
	\item Backtrack matrix: $\Sigma(B)=\{stop,W,S,SW, 1..n\}$
	\ule

\item Initialization: \ul
	\item Cost matrix: $\left\{\begin{array}{l} M_{(i,i)}=0 \\ M_{(i,i-1)}=0 \end{array}\right. \forall i \in 1..{\rm length}(S)$
	\item Backtrack matrix: $\left\{\begin{array}{l} B_{(i,i)}=stop \\ B_{(i,i-1)}=stop \end{array}\right.  \forall i \in 1..{\rm length}(S)$
	\ule

\item Recurrences:
	\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
		M_{(i+1,j-1)}+\omega(i,j) & SW\\
		M_{(i+1,j)} & S\\
		M_{(i,j-1)} & W\\
		\max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} & k
	\end{array}\right\} = B_{(i,j)} \]
	With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.
\item Backtracking: XXX
\item Visualisation: XXX
\item Optimizations: XXX
}
\ole

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Zuker folding}\ol
\item Problem: folding a RNA string $S$ over itself.
{\color{red}
\item Matrices: XXX

\item Alphabets:\ul
\item Input: $\Sigma(S)=\{A,C,G,U\}$.
\item Cost matrix: $\Sigma(W)=\Sigma(V)=\Sigma(F)=\{0..n\},n=$
\item Backtrack matrix: $\Sigma(B)=\{\}$
\ule

\item Initialization:\ul
	\item Cost matrices:\ul
		\item XXX: $W$,$V$
		\item $F_{(0)}=0$
	\ule
	\item Backtrack matrix: $B$
	\ule

\item Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l|l}
	W_{(i+1,j)}+b & S\\
	W_{(i,j-1)}+b & W\\
	V_{(i,j)}+\delta(S_i,S_j) & ?V? \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)} & ?k?
\end{array}\right.\\
\\
V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\
\\
 F_{(j)}&=&\min\left\{\begin{array}{ll} F_{(j-1)} \\ \min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
%
\end{array}\right. \text{(Free Energy)}\\\\
VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
\end{array}\]

\item Backtracking: XXX
\item Visualisation: XXX
\item Optimizations: XXX
}
\ole

%%\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash}m{#1}}
%\newcolumntype{C}{@{\hspace{7pt}}c@{\hspace{7pt}}}
%\def\mnl{\rule{0pt}{2.6ex}\rule[-1.2ex]{0pt}{0pt} \\ \hline}
%$\begin{array}{|C|C|C|C|C|C|} \hline
%0 & 0 & 0 & 0 & 0 & 0 \mnl
%0 &  &  &  &  & \mnl
%0 & M_{23}  &  &  &  & \mnl
%0 &  & \sum  &  &  & \mnl
%0 &  &  &  &  & \mnl
%0 &  &  &  &  & \mnl
%\end{array}$
\end{document}
